Building Sustainable Software - The CSDMS Approach

### 2014-07-14 hutton_et_al_2014_wssspe.tex
Changed name of the tex file for the article.

\author{
  Eric W. H. Hutton$^{\dag}$,
  Mark D. Piper$^{\dag}$,
  Irina Overeem$^{\dag}$
  Albert J. Kettner$^{\dag}$,
}

@@

\csdms{}, The Community Surface Dynamics Modeling System, is an NSF funded
project whose focus is to aid a diverse community of earth and ocean system
model \emph{users} and \emph{developers} to use and create robust software
quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
earth-system models and tools to an international (67 country) community
with the goal of building the set of tools necessary to model the
earth system. \csdms{} is a place modelers look to for access to hundreds open
source surface-dynamics models and tools, as well as model metadata. Such a
model repository increases model transparency and helps to eliminate the
duplication of effors by presenting the current state of modeling efforts.
To increase software sustainability, composability and interoperability \csdms{}
promotes standards that define common modeling interfaces, semantic mediation
between models, and model metadata. Through online resources and workshops,
\csdms{} promotes software engineering best practices, which are foreign to many
developers within our modeling community. For example, version control, unit
testing, continious integration, test-driven development, well-written
\emph{clean code} are all topics of the educational mission of \csdms{}.

@@

\csdms{} has developed innovative model/component interfaces that promote
model reuse and interoperability,
including the Basic Model Interface (\bmi{} - Section~\ref{sec:bmi}), which
uses
the \csdms{} Standard Names (Section~\ref{sec:standardnames}) and a framework
within which models can be coupled and enhanced with a set of \bmi{}-compatible
service components.
The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based client that provides a
graphical interface to this framework and allows users to compose new models
by connecting and configuring components in a simple, browser-based graphical
interface.

@@

\csdms{} has implemented a
Digital Object Identifier (\doi{}) system for software within the model
repository to guarantee recognition to software contributors.
\csdms{} was among the first software venues to assign \doi{}s to
open source software. The advantages of adopting a \doi{} system for open
source software are: 
\begin{itemize}
\item Guarantee credit to the software developer
\item Easily reuse and replicate research as software is directly locatable 
\item Increase software visibility, \doi{} content is 5 times more likely to
      deliver active links than content without.
\item Provide e.g. funding agencies the ability to track usage, so measure
      impact.
\end{itemize}

@@

The \bmi{} identifies the entry points into software components that provides a
calling appication with the necessary level of control to connect components.
\csdms{}, as well as other modeling frameworks such as
ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
OMS~\cite{david2002object}, identifies an interface that, at a minimum,
provides functionality to \emph{initialize}, \emph{update}, and
\emph{finalize} a component. However, for more complicated (and 2-way)
coupling components must implement the \bmi{} methods providing data access,
and model metadata.

@@

%\item \emph{Maintenance}. Because modern browsers tend to adhere to web
%      standards, which lead to fewer cross-compatibility issues than
%      operating systems, only one version of WMT needs to be developed
%      and maintained.

\end{itemize}
With \wmt{}, users can: \emph{run} single component models, build (2-way)
\emph{coupled models}, view and \emph{edit} model parameters, \emph{share}
saved models with others in the community.

@@

\begin{figure}
  \caption{The \csdms{} Web Modeling Tool}
  \begin{center}
    \includegraphics[scale=.25]{wmt.eps}
  \end{center}
  \label{fig:wmt_screenshot}
\end{figure}
\begin{comment}
\subsection{Software sustainability through expanded functionality}

@@

Over the last 4 years over 150 graduate students - from geography, hydrology,
geology, oceanography, earth sciences and engineering departments - have
enrolled in intensive \csdms{}-run modeling courses. These courses
provide an excellent use case for us to test usability and access to the
tool. Students reported in post-course surveys that they were unfamiliar with
the use of models on a high performance computing cluster beforehand (rated
their familiarity as 1.5 average on a scale of 1-5), but ended the courses
with familiarity and being more comfortable with running and changing
simulations as well as visualizing results (average rating approaches 4 on
a Likert scale of 1-5)~\cite{overeem2013strategies}.

@@

\begin{comment}
\huttonnote{Add something about model DOIs in the abstract}

\huttonnote{
\begin{itemize}
\item improving the development process that leads to new software
  \begin{itemize}
  \item methods to develop sustainable software from the outset
  \item effective approaches to reusable software created as a by-product of
        research
  \item impact of computer science research on the development of scientific software
  \end{itemize}
\end{itemize}
}

\huttonnote{This is Irina's section.}

\huttonnote{
\begin{itemize}
\item building large and engaged user communities
  \begin{itemize}
  \item developing strong advocates
  \item measurement of usage and impact
  \end{itemize}
\item encouraging industry's role in sustainability
  \begin{itemize}
  \item engagement of industry with volunteer communities
  \item incentives for industry
  \item incentives for community to contribute to industry-driven projects
  \end{itemize}
\item improving education and training
  \begin{itemize}
  \item best practices for providing graduate students and postdoctoral
        researchers in domain communities with sufficient training in
        software development
  \item novel uses of sustainable software in education (K-20)
  \item case studies from students on issues around software development in
        the undergraduate or graduate curricula
  \end{itemize}
\end{itemize}
}

\huttonnote{This is Albert's section (he's rewriting this)}

\huttonnote{
\begin{itemize}
\item recommending policy changes
  \begin{itemize}
  \item software credit, attribution, incentive, and reward
  \item issues related to multiple organizations and multiple countries, such
        as intellectual property, licensing, etc.
  \item mechanisms and venues for publishing software, and the role of
        publishers
  \end{itemize}
\end{itemize}
}
\end{comment}

>>>>

\author{
  Eric W. H. Hutton$^{\dag}$,
  Mark D. Piper$^{\dag}$,
  Irina Overeem$^{\dag}$,
  Albert J. Kettner$^{\dag}$
}

@@

\csdms{}, The Community Surface Dynamics Modeling System, is an NSF funded
project whose focus is to aid a diverse community of earth and ocean system
model \emph{users} and \emph{developers} to use and create robust software
quickly.  To this end, \csdms{} develops, integrates, archives and disseminates
earth-system models and tools to an international (67 country) community
with the goal of building the set of tools necessary to model the
earth system. Modelers look to \csdms{} for access to hundreds of open
source surface-dynamics models and tools, as well as model metadata. Such a
model repository increases model transparency and helps eliminate
duplication by presenting the current state of modeling efforts.
To increase software sustainability, composability and interoperability,
\csdms{} promotes standards that define common modeling interfaces, semantic
mediation between models, and model metadata. Through online resources and
workshops, \csdms{} promotes software engineering best practices, which are
unfamiliar to many developers within our modeling community. For example,
version control, unit testing, continious integration, test-driven
development, and well-written \emph{clean code} are all topics of the
educational mission of \csdms{}.

@@

\csdms{} has developed innovative model/component interfaces that promote
model reuse and interoperability,
including the Basic Model Interface (\bmi{} - Section~\ref{sec:bmi}), which
uses
the \csdms{} Standard Names (Section~\ref{sec:standardnames}) and a framework
within which models can be coupled and enhanced with a set of \bmi{}-compatible
service components.
The \csdms{} Web Modeling Tool (\wmt{} - Section~\ref{sec:wmt}) is a web-based application that provides a
graphical interface to this framework and allows users to compose new models
by connecting and configuring components in a simple, browser-based graphical
interface.

@@

\csdms{} has implemented a
Digital Object Identifier (\doi{}) system for software within the model
repository to guarantee recognition to software contributors.
\csdms{} was among the first software venues to assign \doi{}s to
open source software. The advantages of adopting a \doi{} system for open
source software are: 
\begin{itemize}
\item Guarantee credit to the software developer
\item Easily reuse and replicate research as software is directly locatable 
\item Increase software visibility: \doi{} content is 5 times more likely to
      deliver active links than content without.
\item Provide funding agencies with the ability to track usage, to measure
      impact.
\end{itemize}

@@

The \bmi{} identifies the entry points into software components that provide a
calling application with the necessary level of control to connect components.
\csdms{}, as well as other modeling frameworks such as
ESMF~\cite{hill2004architecture}, OpenMI~\cite{gregersen2007openmi}, and
OMS~\cite{david2002object}, identifies an interface that, at a minimum,
provides functionality to \emph{initialize}, \emph{update}, and
\emph{finalize} a component. However, for more complicated (and 2-way)
coupling components must implement the \bmi{} methods providing data access,
and model metadata.

@@

\item \emph{Maintenance}. Because modern browsers tend to adhere to web
      standards, which lead to fewer cross-compatibility issues than
      operating systems, only one version of WMT needs to be developed
      and maintained.

\end{itemize}
With \wmt{}, users can: \emph{run} single component models, build (2-way)
\emph{coupled models}, view and \emph{edit} model parameters, and \emph{share}
saved models with others in the community.

@@

\begin{figure}
  \caption{The \csdms{} Web Modeling Tool}
  \begin{center}
    \includegraphics[scale=.25]{wmt.eps}
  \end{center}
  \label{fig:wmt_screenshot}
\end{figure}

\begin{comment}
\subsection{Software sustainability through expanded functionality}

@@

Over the last 4 years over 150 graduate students - from geography, hydrology,
geology, oceanography, earth sciences and engineering departments - have
enrolled in intensive \csdms{}-run modeling courses. These courses
provide an excellent use case for us to test usability and access to our
tools. Students reported in post-course surveys that they were unfamiliar with
the use of models on a high performance computing cluster beforehand (rated
their familiarity as 1.5 average on a scale of 1-5), but ended the courses
with familiarity and being more comfortable with running and changing
simulations as well as visualizing results (average rating approaches 4 on
a Likert scale of 1-5)~\cite{overeem2013strategies}.

@@
